Explore the SingleR results

### 2025-07-19 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/singler-results.Rmd

# only the NBAtlas labels - use for validation dotplot
nbatlas_palette_df <- readr::read_tsv(nbatlas_palette_file)
nbatlas_celltype_pal <- nbatlas_palette_df$hex_color
names(nbatlas_celltype_pal) <- nbatlas_palette_df$NBAtlas_label

@@

## Marker gene expression dotplots

@@

This dotplot shows the top-5 highest log2FC marker genes per NBAtlas cell type for validation.
Marker genes are taken directly from the NBAtlas paper where they were identified with `Seurat::FindMarkers()`.

@@

```{r, fig.width = 32, fig.height = 16}
generate_dotplot(
  merged_sce = merged_sce,
  markers_df = top_nbatlas_markers_df,
  # rename this after we've done all the wrangling
  singler_df = full_celltype_df |> dplyr::rename(label_recoded = label),
  total_cells_df = total_cells_df,
  expressed_genes = expressed_genes,
  bar_order = nbatlas_bar_order, 
  celltype_palette = nbatlas_celltype_pal, 
  min_cells = 0
)
```
For validation, we expect to see high marker gene expression along the diagonal of the dotplot where cell type labels match with their marker genes.
We do broadly see high expression along the diagonal, which means our labeled cells have some amount of similar signal to those in NBAtlas.
Some sets of marker genes show high expression across multiple cell type categories in particular for closely related cell types (e.g. monocyte, macrophage, myeloid, dendritic, or T-cells and NK-cells), but there do not appear to be any strongly unexpected gene expression patterns.

@@

This dotplot shows expression of consensus validation marker genes across matching cell types labeled with `SingleR`.
Harmonized `SingleR` labels with direct matches are shown on the top of the y-axis (`endothelial cell` through `plasma cell`), and the remaining `SingleR` labels are below.

>>>>

# only the NBAtlas labels - use for validation dot plot
nbatlas_palette_df <- readr::read_tsv(nbatlas_palette_file)
nbatlas_celltype_pal <- nbatlas_palette_df$hex_color
names(nbatlas_celltype_pal) <- nbatlas_palette_df$NBAtlas_label

@@

## Marker gene expression dot plots

@@

This dot plot shows the top-5 highest `log2FC` marker genes per NBAtlas cell type for validation.
Marker genes are taken directly from the NBAtlas paper where they were identified with `Seurat::FindMarkers()`.

@@

```{r, fig.width = 32, fig.height = 16}
generate_dotplot(
  merged_sce = merged_sce,
  markers_df = top_nbatlas_markers_df,
  # rename this after we've done all the wrangling
  singler_df = full_celltype_df |> dplyr::rename(label_recoded = label),
  total_cells_df = total_cells_df,
  expressed_genes = expressed_genes,
  bar_order = nbatlas_bar_order, 
  celltype_palette = nbatlas_celltype_pal, 
  min_cells = 0
)
```
For validation, we expect to see high marker gene expression along the diagonal of the dot plot where cell type labels match with their marker genes.
We do broadly see high expression along the diagonal, which means our labeled cells have some amount of similar signal to those in NBAtlas.
Some sets of marker genes show high expression across multiple cell type categories in particular for closely related cell types (e.g. monocyte, macrophage, myeloid, dendritic, or T-cells and NK-cells), but there do not appear to be any strongly unexpected gene expression patterns.

@@

This dot plot shows expression of consensus validation marker genes across matching cell types labeled with `SingleR`.
Harmonized `SingleR` labels with direct matches are shown on the top of the y-axis (`endothelial cell` through `plasma cell`), and the remaining `SingleR` labels are below.
